The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-02-12 paper/paper.tex
adding edge compare
more paper text and content

\section{Results}

@@

@@

@@

@@

Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
and then used classifiers to assign objects to those bins.  Other methods optimized the values
of bin edges themselves.

\begin{itemize}
    \item fisher matrix ellipses overplotted for top N methods
    \item Plots of n(z)
\end{itemize}

\section{Discussion}
It is clear from values above that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration of overall
bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

The inclusion of the $g$ band typically increases metrics by around 10--20\%, including for the
highest scoring methods.  This value should be weighed against the challenge of high-accuracy
PSF measurement for this band, but until the end of the survey this is unlikely to be
the limiting factor in overall precision.

Methods using fixed bins with equal numbers by nominal redshift tend to plateau at scores of
120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
effectively classify by bin, at least in this case of representative, extensive training
information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
training set neighbour to assign bins and thus was explicitly designed to work only
for this idealised scenario; it should give an upper limit when using the same fixed bins.
That other methods with the same nominal bin choice acheive scores close to it once again
confirms that many entries are close to the best possible score for this approach.

Several methods that instead try to optimize nominal bin edges break this
barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The
Stacked Generalization method similarly included manually tuned bin edges to maximize
score for this scenario. The typical 15--20\% improvement in FoM scores when optimizing
nominal bin edges makes clear that this easy win should is compelling.

>>>>

\section{Results \& Discussion}

@@

It is clear from values here that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration of overall
bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

@@

\begin{figure}[htbp]
% \centering
% \includegraphics[width=0.9\linewidth]{results/metric_comparisons.pdf}
\caption{Add $n(z)$ plots here for funbins.}
\label{fig:metrics}
\end{figure}

@@

This is true in particular for the highest scoring methods, so we may assume the behaviour to be
generic. The value of the extra FoM that could be gained should be weighed against 
the challenge of high-accuracy
PSF measurement for this band, and the cost in time and effort of determining it, 
but until the end of the survey this is unlikely to be the limiting factor in overall precision.

@@

Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
and then used classifiers to assign objects to those bins.  Other methods optimized the values
of bin edges themselves, by optimizing the metric.  For the runs shown here, these methods all
targeted one of the 3x2pt metrics. \textbf{Check this is correct}.

A comparison of trained-edge and fixed-edge methods on the CosmoDC2 griz scenarios
is shown in Figure \ref{fig:edges}.  One important point to note is that while trained-edge
methods dominate the highest scores for the 3x2pt DETF metric, on which they trained, they
fare much worse on the lensing-only metrics (the methods were not re-trained on the lensing metrics;
we show the metrics for the same assigned bins).  This suggests that the difference between
optimal bins for different science cases is a significant one, and thus that analysis pipelines
should make it straightforward to calibrate multiple configurations and pass them to
subsequent science anlysis.

\begin{figure*}
\includegraphics[width=0.9\columnwidth]{results/edge_comparison.pdf}
\caption{A comparison of results from methods which used fixed target edges (dashed blue)
with those which optimized target bins (solid orange).  Methods were trained on the 3x2pt metric
shown in the upper panel; the lower performance for these classifications 
for the lensing-only metric shown in the lower panel illustrates the need for science case-specific
binning choices. In each case, classification is done using griz bands on the CosmoDC2 data set.}
\label{fig:edges}
\end{figure*}

Among methods with fixed bins, those using equal numbers in each tend to plateau at scores of
120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
effectively classify by bin, at least in this case of representative, extensive training
information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
training set neighbour to assign bins and thus was explicitly designed to work only
for this idealised scenario; it should give an upper limit when using the same fixed bins.
That other methods with the same nominal bin choice acheive scores close to it once again
confirms that many entries are close to the best possible score for this approach.

Several of the methods with optimized target bin edges break this
barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The
Stacked Generalization method similarly included manually tuned bin edges to maximize
score for this scenario. The typical 15--20\% improvement in FoM scores when optimizing
nominal bin edges makes clear that this easy win should is compelling.
